---
title: "Data Cleaning"
author: "Grace Ramey, Grace.Ramey@ucsf.edu"
output: html_document
### Script to clean the retrospective cohort AID study group
---

A couple of acronyms:
AID = autoimmune disease
Alz or AD = Alzheimer's disease

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(dplyr)
library(DBI)
library(odbc)
library(hash)
library(tibble)
```

```{r load data}
#load the alzheimer's cases and corresponding control data (from alzheimer's queries)
load("prematched_cc_data.Rdata")
#load the autoimmune cases and corresponding control data (from autoimmune queries)
load("prematched_rc_data.Rdata")
#load the data frame that contains the mapping of each autoimmune disease to a disease subtypes
keywords<-readxl::read_xlsx("keywords_2022_10_26.xlsx")
```


Since this is the cleaning script for the retrospective cohort (rc) study group, most of the cleaning will be done on the "prematch_rc" data set:
```{r clean dfs}
#filter out individuals based on lacking demographic data (e.g. gender, ethnicity, etc.)

#missing race data
data_filt <- prematch_rc[!(is.na(prematch_rc$race) | 
                                 prematch_rc$race=="" |
                                 prematch_rc$race=="*Unknown" |
                                 prematch_rc$race=="*Unspecified" |
                                 prematch_rc$race=="Declined" |
                                 prematch_rc$race=="Unknown" |
                                 prematch_rc$race=="Unknown/Declined"),]

#missing ethnicity data
data_filt <- data_filt[!(is.na(data_filt$ethnicity) | 
                                 data_filt$ethnicit=="*Unspecified" |
                                 data_filt$ethnicity=="Declined" |
                                 data_filt$ethnicity=="Unknown" |
                              data_filt$ethnicity=="Unknown/Declined"),]

#missing gender data
data_filt <- data_filt[!(is.na(data_filt$gender) |
               data_filt$gender=="Unknown" |
             data_filt$gender=="*Unspecified"|
             data_filt$gender=="Nonbinary"),]

#*note: for now we will filter out nonbinary gender individuals because of the small sample size of this group

#combine the "Native Hawaiian", "Other Pacific Islander", and "Native Hawaiin or Other Pacific Islander" categories
data_filt$race[data_filt$race == 'Native Hawaiian'] <- "Native Hawaiian or Other Pacific Islander"
data_filt$race[data_filt$race == 'Other Pacific Islander'] <- "Native Hawaiian or Other Pacific Islander"
```


Since patients can have more than one autoimmune disease, they may be duplicated in the data set. Take care of this duplication issue here:
```{r merge duplicated individuals}
#rename data frame (also do this so you have a copy)
aid<-data_filt
#remove individual disease columns (will reassign after grouping)
aid<-select(aid, -keywords$Col_names)

#add some new columns that will be filled in the for loop
aid<-aid%>%add_column(all_AIDs=NA) %>% add_column(aid_dates=NA) %>% 
  add_column(earliest_aid_date=NA) %>% add_column(earliest_aid=NA)
#group by person id and merge unique AID names for each person into a single string:
for(i in 1:length(aid$person_id)){
  if(i%%40000==0){print(i)}
  if(aid$aid[i]==1){
    #get the current person id
    cur_id<-aid$person_id[i]
    #filter the data frame to only include the current person id
    aid_filt<-aid %>% filter(person_id==cur_id)
    #collapse the unique autoimmune disease names for the current person into a single string
    all_aids<-paste(aid_filt$AID_name, collapse=", ")
    #collapse the unique autoimmune disease dx dates for the current person into single string
    all_times<-paste(aid_filt$aid_date, collapse=", ")
    #add these values to the new columns
    aid$all_AIDs[i]<-all_aids
    aid$aid_dates[i]<-all_times
    #get the minimum aid dx date
    aid$earliest_aid_date[i]<-min(aid_filt$aid_date)
    ead<-aid$earliest_aid_date[i]
    #get the name of the earliest autoimmune disease that occurred in the patient
    #(this step will be most important for patients with multiple autoimmune diseases)
    aid$earliest_aid[i]<- paste((aid_filt %>% filter(aid_date == ead))$AID_name,
                                collapse=", ")
  }
}

#now that all rows for a patient should contain the same data...remove duplicated person ids
aid_filt<-select(aid, -c(AID_name, aid_date))
aid_distinct<-distinct(aid_filt)
print(length(aid_distinct$person_id) == length(unique(aid_distinct$person_id)))

#Great! Now there are no duplicates

#add back in the columns for individual diseases
n_col<-ncol(aid_distinct)
for(i in 1:length(keywords$AID_name)){
  aid_distinct<-aid_distinct %>% add_column(temp=0)
  colnames(aid_distinct)[colnames(aid_distinct) == 'temp'] <- keywords$Col_names[i]
}

#add 1s to the aid columns depending on the aid names for each person
for(i in 1:length(aid_distinct$person_id)){
  if(aid_distinct$aid[i]==1){
    for(j in 1:length(keywords$AID_name)){
      if(grepl(aid_distinct$all_AIDs[i], pattern=keywords$AID_name[j])){
        aid_distinct[[n_col+j]][i]<-1
      }
    }
  }
}

#calculate the number of autoimmune diseases each person has
aid_distinct<-aid_distinct %>% add_column(numAIDs = 0)
for(i in 1:length(aid_distinct$person_id)){
  aid_distinct$numAIDs[i]<-rowSums(aid_distinct[i, (n_col+1):(ncol(aid_distinct))], na.rm=TRUE)
}

#change the earliest_aid date column to a date object
aid_distinct$earliest_aid_date<-as.Date(aid_distinct$earliest_aid_date, origin='1970-01-01')

```


```{r add alz indication}
#indicate if each aid_distinct person has alzheimer's
#do this by intersecting those with alz from the cc data set with the aid_distinct data set
alz_df <- prematch_cc %>% filter(alz==1)
aid_distinct<-aid_distinct %>% add_column(alz=0) %>% add_column(alz_date=NA)
for(i in 1:length(aid_distinct$person_id)){
  #get the current person id
  cur_id<-aid_distinct$person_id[i]
  if(cur_id %in% alz_df$person_id){
    #if the person_id is in the alz data frame, add a one to that person's alz column
    aid_distinct$alz[i]<-1
    alz_filt<-alz_df %>% filter(person_id==cur_id)
    #also get the date when the person was diagnosed with alz
    aid_distinct$alz_date[i]<-alz_filt$alz_date[1]
  }
}

#change the alz_date column to a date object
aid_distinct$alz_date<-as.Date(aid_distinct$alz_date, origin='1970-01-01')
```


```{r disease subtypes/groupings}
#add an indication for which subtype group the autoimmune diseases fall into
col_n<- ncol(aid_distinct) #get current number of columns in the data frame
#add columns to hold 1s or 0s for each disease subtype group for each patient
for(i in 1:length(unique(keywords$Group))){
  aid_distinct<-aid_distinct %>% add_column(temp = 0)
  colnames(aid_distinct)[colnames(aid_distinct) == 'temp'] <- unique(keywords$Group)[i]
}

#add a column to hold the name(s) of the disease subtype(s) a patient has
aid_distinct<- aid_distinct %>% add_column(disease_groups = "")

#create a hash to lookup the group based on each AID name
group_hash<-hash()
for(i in 1:length(keywords$AID_name)){
  group_hash[[keywords$AID_name[i]]]<-keywords$Group[i]
}

#now assign each patient's autoimmune disease(s) to a disease group(s)
for(i in 1:length(aid_distinct$all_AIDs)){
  group_list<-c() #vector that will hold disease group names for each person
  for(j in 1:length(keywords$AID_name)){
    #if the current disease name is in the patient's autoimmune disease list...
    if(grepl(aid_distinct$all_AIDs[i], pattern=keywords$AID_name[j])){
      #look up the subtype group the disease name falls under
      group_name<-group_hash[[keywords$AID_name[j]]]
      #add a 1 to the column representing this group
      aid_distinct[[group_name]][i] <- 1
      group_list<-c(group_list, group_name)
    }
  }
  group_list<-unique(group_list)
  aid_distinct$disease_groups[i]<-paste(group_list, collapse=", ")
  if(i%%40000==0){print(i)} #print a checkpoint
}
```


We are trying to investigate the effect that exposure to autoimmune diseases has on Alzheimer's patients, so remove individuals who had an autoimmune disease diagnosed *after* Alzheimer's:
```{r Alz before AID?--if so, remove}
#remove individuals that were diagnosed with alzheimer's before their autoimmune condition, to eliminate reverse causality as much as possible
aid_distinct<-aid_distinct %>% add_column(remove=0)
for(i in 1:length(aid_distinct$person_id)){
  if(aid_distinct$aid[i]==1 && aid_distinct$alz[i]==1){
    if(aid_distinct$earliest_aid_date[i]>aid_distinct$alz_date[i]){
      aid_distinct$remove[i]<-1
    }
  }
}

#check how many patients this step will remove
sum(aid_distinct$remove)
#filter out the individuals
aid_distinct<-aid_distinct %>% filter(remove==0)
aid_distinct<-select(aid_distinct, -remove)
```



```{r calculate some other important metrics for matching and quality control}
#calculate the length of the person's ehr record (useful for matching)
aid_distinct<- aid_distinct %>% add_column(ehr_time_days = NA)
for(i in 1:length(aid_distinct$person_id)){
  aid_distinct$ehr_time_days[i]<-aid_distinct$last_visit[i] - aid_distinct$first_visit[i]
}

#filter out anyone with 0 visits or ehr length of 0 days
aid_distinct<- aid_distinct %>% add_column(remove = 0)
for(i in 1:length(aid_distinct$person_id)){
  if(aid_distinct$ehr_time_days[i]==0){aid_distinct$remove[i]<-1}
  if(aid_distinct$num_visits[i]==0){aid_distinct$remove[i]<-1}
}
sum(aid_distinct$remove)
aid_distinct<-aid_distinct %>% filter(remove==0)
aid_distinct<-select(aid_distinct, -remove)

#calculate age at AD diagnosis
aid_distinct<-aid_distinct %>% add_column(aaAD_years = NA)
for(i in 1:length(aid_distinct$person_id)){
  #print(i)
  if(aid_distinct$alz[i]==1){
    #age at AD diagnosis
    aid_distinct$aaAD_years[i]<-as.numeric(format(as.Date(as.character(aid_distinct$alz_date[i]),
                                                    origin='1970-01-01'), format="%Y")) -
                                                      aid_distinct$birth_year[i]
  }
}
```

```{r calculate age at death}
#calculate age at death so you can match/filter people by lifespan
aid_distinct<-aid_distinct%>%add_column(aaDeath_years=NA)
for(i in 1:length(aid_distinct$person_id)){
  if(i%%40000==0){print(i)}
  #age at death
  ##replace 'redacted' column with column that contains patient death date information 
  aid_distinct$aaDeath_years[i]<-as.numeric(format(as.Date(
    as.character(aid_distinct$redacted[i]),
                                                  origin='1970-01-01'), format="%Y")) -
                                                    aid_distinct$birth_year[i]
}

```

```{r finally, save the cleaned data}
#save this final data frame
save(aid_distinct, file='M:/Prematch_data/prematch_rc_clean.Rdata')
```