---
title: "Data Matching"
author: "Grace Ramey, Grace.Ramey@ucsf.edu"
output: html_document
---

This notebook is meant for matching autoimmune disease individuals to non-autoimmune controls (for the retrospective cohort group), and alzheimer's individuals with non-alzheimer's controls (for the case-control group).


```{r load data and libraries}
#load data sets that contain cases and controls
load('M:/Prematch_data/prematch_cc_clean.Rdata') #case-control (cc) study group
load('M:/Prematch_data/prematch_rc_clean.Rdata') #retrospective cohort (rc) study group
cc<-alz
rc<-aid_distinct
rm(alz, aid_distinct)
keywords<-readxl::read_xlsx("~/EHR/keywords_2022_10_26.xlsx") #keywords data frame
library(dplyr)
library(MatchIt)
library(hash)
library(tibble)
```


```{r standardize before matching}
#convert certain variables to factors
cc$race <- as.factor(cc$race)
cc$ethnicity <- as.factor(cc$ethnicity)
cc$gender <- as.factor(cc$gender)
cc$birth_year<-as.factor(as.character(cc$birth_year))

rc$race <- as.factor(rc$race)
rc$ethnicity <- as.factor(rc$ethnicity)
rc$gender <- as.factor(rc$gender)
rc$birth_year<-as.factor(as.character(rc$birth_year))

#standardize column names between cc and rc
colnames(cc)[colnames(cc) == 'AID_name'] <- "all_AIDs"
colnames(cc)[colnames(cc) == 'earliest_aid_time'] <- "earliest_aid_date"
```


```{r cc matching using propensity scores}
## can impose some age filters here prior to matching
#cc<-cc %>% filter(aaDeath_years>=30) %>% filter(aaDeath_years<100)
#cc<-cc %>% filter(aaDeath_years>=85) %>% filter(aaDeath_years<100)
#cc<-cc %>% filter(aaDeath_years>=80) %>% filter(aaDeath_years<100)

#create a matching object for the case-control data set
#can add or remove variables for different sensitivity analyses
m.out1 <- matchit(alz ~ birth_year + gender + race + ethnicity + aaDeath_years,
                  data = cc,
                  method='nearest',
                  exact=c('gender', 'race', 'ethnicity'),
                  ratio=1,
                  verbose=T)
#assess quality of the match
plot(summary(m.out1))

#extract the matched data set from the object
cc_matched<-match.data(m.out1, distance='prop.score')
#save the matched data
save(cc_matched,m.out1, file='M:/Matched_data/matched_cc_clean_m1.Rdata')

#Note: I use the m1/m2/m3/... etc. notation to denote which kind of match criteria I used for a particular data set

#e.g., m1 means I matched on demographic variables (birth year, gender, ethnicity, and race) as well as age at death, whereas m3 means I only matched on birth year, gender, ethnicity, and race
```


```{r final cc quality control steps}
#for cc cohort...

#mark pairs for removal if either person had alz diagnosed before study enrollment
#here, we consider study enrollment to be the date when a person was diagnosed with an autoimmune disease
cc_matched<- cc_matched %>% add_column(remove=0)
cc_matched<-cc_matched %>% arrange(subclass) #sort the dataset by subclass
rownames(cc_matched) <- 1:nrow(cc_matched) #re-index the ordered dataframe
for(i in 1:length(cc_matched$person_id)){
  if((i%%2) == 0){
    #copy over the alz date of the case to the control
    cc_matched$alz_date[i]<-cc_matched$alz_date[i-1]
  }
}
#if the control person has an aid before the alz date, mark the entire pair for removal
for(i in 1:length(cc_matched$person_id)){
  if((i%%2) == 0){
    if(cc_matched$aid[i]==1 && (cc_matched$earliest_aid_date[i] > cc_matched$alz_date[i])){
      #mark both individuals in the pair for removal
      cc_matched$remove[i]<-1
      cc_matched$remove[i-1]<-1
    }
  }
}
sum(cc_matched$remove)
cc_matched<-cc_matched %>% filter(remove==0)
cc_matched<-select(cc_matched, -remove)

#reset the subclass to start from 1 since certain subclasses were filtered out
cc_matched<-cc_matched %>% add_column(new_sc=0)
count<-1
for(i in 1:length(cc_matched$person_id)){
  if(i%%2==0){
    cc_matched$new_sc[i]<-cc_matched$new_sc[i-1]
  }
  else{
    cc_matched$new_sc[i]<-count
    count<-count+1
  }
}
cc_matched<-select(cc_matched,-subclass)
colnames(cc_matched)[colnames(cc_matched) == 'new_sc'] <- "subclass"

#save
save(cc_matched, m.out1, file='M:/Matched_data/matched_cc_clean_m1.Rdata')
```



```{r rc matching using propensity scores}
## can impose some age filters here
#rc<-rc %>% filter(aaDeath_years>=30) %>% filter(aaDeath_years<100)
#rc<-rc %>% filter(aaDeath_years>=30) %>% filter(aaDeath_years<100)
#rc<-rc %>% filter(aaDeath_years>=65) %>% filter(aaDeath_years<100)
#rc<-rc %>% filter(aaDeath_years>=80) %>% filter(aaDeath_years<100)

#create a matching object for the retrospective cohort data set
#removed aaDeath_years as a matching variable
m.out2 <- matchit(aid ~ birth_year + gender + race + 
                    ethnicity + aaDeath_years,
                  data = rc,
                  method='nearest',
                  exact=c('gender', 'race', 'ethnicity'),
                  ratio=1,
                  verbose=T)
#assess quality of the match
plot(summary(m.out2))

#extract the matched data set from the object
rc_matched<-match.data(m.out2, distance='prop.score')
#save the matched data
save(rc_matched, m.out2, file='M:/Matched_data/matched_rc_clean_m1.Rdata')
```



```{r Final rc quality control steps}
#for rc cohort...

#assign study enrollment dates to the controls based on the aid diagnosis date of the case they were matched with
m.data1_copy<-rc_matched
rc_matched<-rc_matched %>% arrange(subclass) #sort the dataset by subclass
rownames(rc_matched) <- 1:nrow(rc_matched) #re-index the ordered dataframe
n_diseases<-length(keywords$AID_name) #establish some useful numbers for the for loop
n_groups<-length(unique(keywords$Group)) #establish some useful numbers for the for loop
earliest_aid_idx<-which(colnames(rc_matched)=="earliest_aid") #get the "alz" column number
alz_time_idx<-which(colnames(rc_matched)=="alz_date") #get the number of the column "alz_date"
for(i in 1:length(rc_matched$person_id)){
  #even indices are the controls
  if((i%%2) == 0){
    #copy over the disease group from case to control
    rc_matched$disease_groups[i]<-rc_matched$disease_groups[i-1]
    #copy aid_names
    rc_matched$all_AIDs[i]<-rc_matched$all_AIDs[i-1]
    #copy earliest aid and earliest aid date
    rc_matched$earliest_aid_date[i]<-rc_matched$earliest_aid_date[i-1]
    rc_matched$earliest_aid[i]<-rc_matched$earliest_aid[i-1]
  }
  else{
    #copy 1s and 0s into disease columns
    for(j in 1:n_diseases){
      if(rc_matched[[colnames(rc_matched)[earliest_aid_idx+j]]][i] == 1){
        rc_matched[[colnames(rc_matched)[earliest_aid_idx+j]]][i+1]<-1
      }
    }
    
    for(k in 1:n_groups){
      #copy 1s and 0s into disease group columns
      if(rc_matched[[colnames(rc_matched)[alz_time_idx+k]]][i] == 1){
        rc_matched[[colnames(rc_matched)[alz_time_idx+k]]][i+1]<-1
      }
    }
  }
}


#mark pairs for removal if they had alz diagnosed before study enrollment
rc_matched<- rc_matched %>% add_column(remove=0)
for(i in 1:length(rc_matched$person_id)){
  if((i%%2) == 0){
    if(rc_matched$alz[i]==1){
      if(rc_matched$earliest_aid_date[i]>rc_matched$alz_date[i]){
        #mark both individuals in the pair for removal
        rc_matched$remove[i]<-1
        rc_matched$remove[i-1]<-1
      }
    }
  }
}
sum(rc_matched$remove)
rc_matched<-rc_matched %>% filter(remove==0)
rc_matched<-select(rc_matched, -remove)

#reset the subclass to start from 1 since certain subclasses were filtered out
rc_matched<-rc_matched %>% add_column(new_sc=0)
count<-1
for(i in 1:length(rc_matched$person_id)){
  if(i%%2==0){
    rc_matched$new_sc[i]<-rc_matched$new_sc[i-1]
  }
  else{
    rc_matched$new_sc[i]<-count
    count<-count+1
  }
}
rc_matched<-select(rc_matched,-subclass)
colnames(rc_matched)[colnames(rc_matched) == 'new_sc'] <- "subclass"

#save the data
save(rc_matched, m.out2, file='M:/Matched_data/matched_rc_clean_m1.Rdata')
```


Temporal data matching for the longitudinal age at AD onset analysis. Take the AD cases from the aid cohort, and match them to AD cases from the background control cohort
```{r}
#get people with autoimmune diseases who also had alzheimer's
ad_aid<-rc %>% filter(aid==1) %>% filter(alz==1)
#get people without autoimmune diseases who also had alzheimer's
ad_noaid<-rc %>% filter(aid==0) %>% filter(alz==1)
#bind together
ad<-rbind(ad_aid, ad_noaid)

m.out3<- matchit(aid ~ birth_year + gender + race + ethnicity + aaDeath_years,
                  data = ad, 
                 discard = "both", #May or may not need to keep this (keep if the match is poor without it, as this will discard poorly matched individuals)
                 exact = c('gender', 'race', 'ethnicity'),
                  method='nearest',
                  ratio=1,
                  verbose=T)
#check quality of the match
plot(summary(m.out3))
#get data from the matchit object
rc_temporal<-match.data(m.out3, distance='prop.score')
#save the matched data
save(rc_temporal,m.out3, file='M:/Matched_data/matched_rc_temporal.Rdata')
```


```{r final temporal cohort quality control}
#assign study enrollment dates to the controls based on the aid diagnosis date of the case they were matched with
m.data3_copy<-rc_temporal
rc_temporal<-rc_temporal %>% arrange(subclass) #sort the dataset by subclass
rownames(rc_temporal) <- 1:nrow(rc_temporal) #re-index the ordered dataframe
n_diseases<-length(keywords$AID_name) #establish some useful numbers for the for loop
n_groups<-length(unique(keywords$Group)) #establish some useful numbers for the for loop
earliest_aid_idx<-which(colnames(rc_temporal)=="earliest_aid") #get the "alz" column number
alz_time_idx<-which(colnames(rc_temporal)=="alz_date") #get the number of the column "alz_date"
for(i in 1:length(rc_temporal$person_id)){
  #even indices are the controls
  if((i%%2) == 0){
    #copy over the disease group from case to control
    rc_temporal$disease_groups[i]<-rc_temporal$disease_groups[i-1]
    #copy aid_names
    rc_temporal$all_AIDs[i]<-rc_temporal$all_AIDs[i-1]
    #copy earliest aid and earliest aid date
    rc_temporal$earliest_aid_date[i]<-rc_temporal$earliest_aid_date[i-1]
    rc_temporal$earliest_aid[i]<-rc_temporal$earliest_aid[i-1]
  }
  else{
    #copy 1s and 0s into disease columns
    for(j in 1:n_diseases){
      if(rc_temporal[[colnames(rc_temporal)[earliest_aid_idx+j]]][i] == 1){
        rc_temporal[[colnames(rc_temporal)[earliest_aid_idx+j]]][i+1]<-1
      }
    }
    
    for(k in 1:n_groups){
      #copy 1s and 0s into disease group columns
      if(rc_temporal[[colnames(rc_temporal)[alz_time_idx+k]]][i] == 1){
        rc_temporal[[colnames(rc_temporal)[alz_time_idx+k]]][i+1]<-1
      }
    }
  }
}


#mark pairs for removal if they had alz diagnosed before study enrollment
rc_temporal<- rc_temporal %>% add_column(remove=0)
for(i in 1:length(rc_temporal$person_id)){
  if((i%%2) == 0){
    if(rc_temporal$alz[i]==1){
      if(rc_temporal$earliest_aid_date[i]>rc_temporal$alz_date[i]){
        #mark both individuals in the pair for removal
        rc_temporal$remove[i]<-1
        rc_temporal$remove[i-1]<-1
      }
    }
  }
}
sum(rc_temporal$remove)
rc_temporal<-rc_temporal %>% filter(remove==0)
rc_temporal<-select(rc_temporal, -remove)

#reset the subclass to start from 1 since certain subclasses were filtered out
rc_temporal<-rc_temporal %>% add_column(new_sc=0)
count<-1
for(i in 1:length(rc_temporal$person_id)){
  if(i%%2==0){
    rc_temporal$new_sc[i]<-rc_temporal$new_sc[i-1]
  }
  else{
    rc_temporal$new_sc[i]<-count
    count<-count+1
  }
}
rc_temporal<-select(rc_temporal,-subclass)
colnames(rc_temporal)[colnames(rc_temporal) == 'new_sc'] <- "subclass"

#save the data
save(rc_temporal, m.out3, file='M:/Matched_data/matched_rc_temporal.Rdata')
```
